---
title: "`r paste(params$name)` (baconed)"
subtitle: "`r paste(params$covariates)`"
date: "`r format(Sys.time(), '%d-%m-%Y')`"
params:

# The bit on the RIGHT needs to be exactly what is defined in the render function, left is need for the rmd

  name:               name
  raw_results:        raw_results
  covariates:         covariate_names
  results_summary:    results_summary
  date_started:       date_started
  p_data:             p
  voi:                var 
  
output: 
  html_document:
    number_sections: true
    fig_caption: true
    df_print: paged
    toc: true
    toc_float: true
    theme: flatly
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#setwd("E:/BMIQ MAV/19-07-30-split_cohort_and_run_season/reports/tG_summer")
knitr::opts_knit$set(root.dir = here::here())
options("scipen"=2, "digits"=3)
```

```{r libraries, include=FALSE}
#library(GenABEL)
# library(missMethyl)
# library(limma) 
# library(IlluminaHumanMethylationEPICanno.ilm10b4.hg19)
library(tidyverse)
library(kableExtra)
library(formattable)
library(Hmisc)
library(ggpubr)
library(gridExtra)
library(rstatix)
library(moments)
```

```{r set-up directories, include=FALSE}

#dir.create(paste0("./reports/", "outs/", params$name), recursive = T)
#dir.create(paste0("./reports/", "universes"), recursive = T)

p   <- params$p_data
voi <- params$voi
#data_output <-      paste0("./reports", "/", params$name, "/")
var_raw <-params$raw_results
var<-var_raw

# ----------------------------------------------------------

#colnames(var)<-cols
var <- var %>% arrange(Bacon_FDR, Bacon_P) %>% rowid_to_column("Rank")
var$ucsc_refgene_name<-sub(";.*", "", var$ucsc_refgene_name)
var$ucsc_refgene_group<-sub(";.*", "", var$ucsc_refgene_group)

# Define functions
top_hits <- function(that){
  output<- vector("character", nrow(that))
  for(i in 1:nrow(that)){
    if((is.na(that$ucsc_refgene_name[[i]]) | that$ucsc_refgene_name[[i]] == "") == TRUE){
      output[[i]] <- that$row.names[[i]]
    }
    else{
      output[[i]] <- that$ucsc_refgene_name[[i]]
    }
  }
  output <- enframe(output, name = NULL, value = "Top 100 hits")
  output <- dplyr::slice(output, 1:100)
  return(output)
}

cont_var_normality <- function(df, variable){

  df <- df[!is.na(df[[variable]]), ]
  p1 <- ggdensity(df[[variable]], fill = "lightgray") +
    stat_overlay_normal_density(color = "red", linetype = "dashed")
  p2 <- ggplot(df, aes(x = df[[variable]])) +
    geom_histogram(color = "darkblue", fill = "lightblue")
  p3 <- ggqqplot(df[[variable]])
  p4 <- ggplot(df, aes(x = "", y = df[[variable]])) +
    geom_boxplot(outlier.colour = "red",
                 outlier.size = 2)
  #print(paste0("Number of values for ", variable, " = ", length(df[[variable]])))
  grid.arrange(p1, p2, p3, p4, ncol = 2)
  #grid.arrange(p2, p4, ncol = 2)
}

counting <- function(df, voi){
  orbs <- sym(paste0(voi))
  df %>% count({{orbs}})
}

gg_qqplot <- function(var_ps, ci = 0.95) {
  n  <- length(var_ps)
  df <- data.frame(
    observed = -log10(sort(var_ps)),
    expected = -log10(ppoints(n)),
    clower   = -log10(qbeta(p = (1 - ci) / 2, shape1 = 1:n, shape2 = n:1)),
    cupper   = -log10(qbeta(p = (1 + ci) / 2, shape1 = 1:n, shape2 = n:1))
  )
  log10Pe <- expression(paste("Expected -log"[10], plain(P)))
  log10Po <- expression(paste("Observed -log"[10], plain(P)))
  ggplot(df) +
    geom_point(aes(expected, observed), shape = 1, size = 3) +
    geom_abline(intercept = 0, slope = 1, alpha = 0.5) +
    geom_line(aes(expected, cupper), linetype = 2) +
    geom_line(aes(expected, clower), linetype = 2) +
    xlab(log10Pe) +
    ylab(log10Po) +
    ggtitle("qqplot")
}
```

# Distribution / breakdown of `r voi`

```{r, message=FALSE, warning=FALSE, echo=FALSE}
if(!is.factor(p[[voi]])){
cont_var_normality(p, voi)
} else {
  counting(p, voi)
}
```

<br />

# Lambdas & covariates {.tabset .tabset-fade .tabset-pills}

## Lambdas

```{r sums and lambdas, echo=FALSE}

rl <- params$results_summary
rl 

```

<br />


# Top 500 array hits

```{r top 500 genes, echo=FALSE, warning=FALSE}
#head(var)

var %>%
  mutate_at(c("logFC",
              "AveExpr",
              "t",
              "P.Value",
              "adj.P.Val",
              "Bacon_P",
              "Bacon_FDR",
              "B"
              ), ~signif(., 3)) %>%
  arrange(Bacon_FDR, Bacon_P) %>%
  dplyr::slice(1:500) %>%
  DT::datatable(
    filter = 'top',
    rownames = FALSE,
    extensions = list('FixedColumns' = NULL,
                      'Buttons' = NULL),
    options = list(pageLength = 15,
                   dom = 'Bfrtip',
                   scrollX = TRUE,
                   fixedColumns = TRUE,
                   columnDefs = list(
                     list(
                       className = 'dt-center', 
                       targets = 0:(ncol(var)-1))
                     ),
                   buttons = list(
                       'copy',
                     list(
                       extend = 'colvisGroup',
                       text = 'simplified',
                       show = c(0,1,3,9,10),
                       hide = c(2,4,5,6,7,8,11,12,13,14,15,16,17,18)
                       ),
                     list(
                       extend = 'colvisGroup',
                       text = 'show all',
                       show = c(0:18)
                       ),
                     list(
                       extend = 'colvis',
                       text = 'see columns',
                       collectionLayout= 'fixed two-column'
                       )
                     )
    )
  ) %>%
  #DT::formatRound(c('logFC', 'estimate', 't', 'P.Value', 'adj.P.Val', 'B'), 5) %>%
  DT::formatStyle(columns = c(1:14, 16), `text-align` = 'center') %>%
  DT::formatStyle(columns = c(18,19), `text-align` = 'left') %>%
  DT::formatStyle(
    'Bacon_FDR',
    #color = DT::styleInterval(0.25, c('black', 'Black')),
    backgroundColor = DT::styleInterval(0.25, c('#b3de69', '')),
    fontWeight = DT::styleInterval(0.25, c('bold', ''))
  )
  # DT::formatStyle(
  #   'estimate',
  #   backgroundColor = DT::styleInterval(0.80, c('', '#fbb4ae')),
  #   fontWeight = DT::styleInterval(0.80, c('', 'bold'))
  # )
```

<br />

# CpG subsets {.tabset .tabset-fade .tabset-pills}

<br />

## Top 100 hits

```{r, include=FALSE}
totty <- top_hits(var)
```

```{r, echo = FALSE}
totty %>%
  #dplyr::rename("Top 100 hits" = "") %>% 
  #mutate_at(7, funs(round(., 3))) %>%
  DT::datatable(
    filter = 'top',
    rownames = FALSE,
    extensions = list('FixedColumns' = NULL,
                      'Buttons' = NULL),
    options = list(pageLength = 15,
                   dom = 'Brtip',
                   fixedColumns = TRUE,
                   columnDefs = list(
                     list(
                       className = 'dt-center', 
                       targets = "_all")
                   ),
                   buttons = 'copy'
    )
  )
```

<br />

## Top 100 dmCpG-associated genes

```{r top 100 named genes, include=FALSE}
named_hits <- var[!(is.na(var$ucsc_refgene_name) | var$ucsc_refgene_name == ""),]
top100_named_genes <-
  named_hits %>%
  arrange(Bacon_FDR, Bacon_P) %>%
  dplyr::slice(1:100) %>%
  dplyr::select(ucsc_refgene_name)
```

```{r top 100 named genes graph, echo=FALSE}
top100_named_genes %>%
  dplyr::rename("Top 100 dmCpG-associated genes" = "ucsc_refgene_name") %>% 
  #mutate_at(7, funs(round(., 3))) %>%
  DT::datatable(
    filter = 'top',
    rownames = FALSE,
    extensions = list('FixedColumns' = NULL,
                      'Buttons' = NULL),
    options = list(pageLength = 15,
                   dom = 'Brtip',
                   fixedColumns = TRUE,
                   columnDefs = list(
                     list(
                       className = 'dt-center', 
                       targets = "_all")
                   ),
                   buttons = 'copy'
                   )
    )
```

<br />

## Top 100 gene-associated dmCpGs 

```{r top 100 cgs, include=FALSE}
top100_cgs <- 
  named_hits %>%
  arrange(Bacon_FDR, Bacon_P) %>%
  dplyr::slice(1:100) %>%
  dplyr::select(row.names) 
```

```{r top 100 cgs graph, echo=FALSE}
top100_cgs %>%
  dplyr::rename("Top 100 gene-associated dmCpGs" = "row.names") %>% 
  #mutate_at(7, funs(round(., 3))) %>%
  DT::datatable(
    filter = 'top',
    rownames = FALSE,
    extensions = list('FixedColumns' = NULL,
                      'Buttons' = NULL),
    options = list(pageLength = 15,
                   dom = 'Brtip',
                   fixedColumns = TRUE,
                   columnDefs = list(
                     list(
                       className = 'dt-center', 
                       targets = "_all")
                   ),
                   buttons = 'copy'
                   )
    )
```

<br />

## Significant dmCpG-associated genes

```{r top sig genenames, include=FALSE}
named_hits <- var[!(is.na(var$ucsc_refgene_name) | var$ucsc_refgene_name == ""),]
sig_genenames <-
  named_hits %>%
  dplyr::arrange(Bacon_FDR, Bacon_P) %>%
  dplyr::filter(Bacon_FDR <0.25) %>%
  dplyr::select(ucsc_refgene_name)
```

```{r top sig genenames table, echo=FALSE}
sig_genenames %>%
  dplyr::rename("Significant hits associated with named genes" = "ucsc_refgene_name") %>% 
  #mutate_at(7, funs(round(., 3))) %>%
  DT::datatable(
    filter = 'top',
    rownames = FALSE,
    extensions = list('FixedColumns' = NULL,
                      'Buttons' = NULL),
    options = list(pageLength = 15,
                   dom = 'Brtip',
                   fixedColumns = TRUE,
                   columnDefs = list(
                     list(
                       className = 'dt-center', 
                       targets = "_all")
                   ),
                   buttons = 'copy'
                   )
    )
```

<br />

## Significant gene-associated dmCpGs 

```{r top sig cgs, include=FALSE}
sig_cgs <-
  var %>%
  dplyr::arrange(Bacon_FDR, Bacon_P) %>%
  dplyr::filter(Bacon_FDR <0.25) %>%
  dplyr::select(row.names)
```

```{r top sig cgs graph, echo=FALSE}
sig_cgs %>%
  dplyr::rename("Significant cgs" = "row.names") %>% 
  DT::datatable(
    filter = 'top',
    rownames = FALSE,
    extensions = list('FixedColumns' = NULL,
                      'Buttons' = NULL),
    options = list(pageLength = 15,
                   dom = 'Brtip',
                   fixedColumns = TRUE,
                   columnDefs = list(
                     list(
                       className = 'dt-center', 
                       targets = "_all")
                   ),
                   buttons = 'copy'
                   )
    )
```

<br />


# Graphs {.tabset .tabset-fade .tabset-pills}

## Report qqplot and nominal p.values {.tabset .tabset-fade .tabset-pills}

### qqplot - pre-bacon

```{r, include=FALSE, warning=FALSE}
var_ps <- var$P.Value
```

```{r, echo=FALSE}
gg_qqplot(var_ps)
```

### qqplot - post-bacon

```{r, include=FALSE, warning=FALSE}
var_ps <- var$Bacon_P
```

```{r, echo=FALSE}
gg_qqplot(var_ps)
```

### p.values - pre-bacon

```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(var, aes(x=var$P.Value)) + 
  geom_histogram(color="darkblue", fill="lightblue", boundary = 0) +
  ggtitle("P-values: var")
```

### p.values - post_bacon

```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(var, aes(x=var$Bacon_P)) + 
  geom_histogram(color="darkblue", fill="lightblue", boundary = 0) +
  ggtitle("P-values: var")
```


## CpG location data {.tabset .tabset-fade .tabset-pills}

```{r pie chart data, include=FALSE}
## Relation to island 
pie_rel_to_island_all <-
  var %>% 
  dplyr::count(relation_to_island) %>% 
  dplyr::arrange(desc(n)) 

pie_region_100 <-
  var[!(is.na(var$relation_to_island) | var$relation_to_island == ""),]
pie_rel_to_island_100 <-
  pie_region_100 %>%
  arrange(Bacon_FDR, Bacon_P) %>%
  dplyr::slice(1:100) %>%
  dplyr::select(relation_to_island) %>%
  dplyr::count(relation_to_island) %>%
  arrange(desc(n))
  #rename(relation_to_island_100 = relation_to_island) %>% 


## RefGene group
pie_ref_group_all <-
  var %>% 
  dplyr::count(ucsc_refgene_group) %>% 
  dplyr::arrange(desc(n))
pie_ref_group_all$ucsc_refgene_group<-sub("^$", "Unknown", pie_ref_group_all$ucsc_refgene_group)

pie_region_100 <-
  var[!(is.na(var$ucsc_refgene_group) | var$ucsc_refgene_group == ""),]
pie_ref_group_100 <-
  pie_region_100 %>%
  dplyr::arrange(Bacon_FDR, Bacon_P) %>%
  dplyr::slice(1:100) %>%
  dplyr::select(ucsc_refgene_group) %>%
  dplyr::count(ucsc_refgene_group) %>% 
  dplyr::arrange(desc(n)) 
  #rename(ucsc_refgene_group_100 = ucsc_refgene_group) %>% 
pie_ref_group_100$ucsc_refgene_group<-sub("^$", "Unknown", pie_ref_group_100$ucsc_refgene_group)
```

### Relationship to island

```{r, fig.cap="Relationship to island of the top 100 most significant array hits (*left*) and all hits (right)", echo = FALSE}
colours <- list('#b3de69','#8dd3c7','#ffffb3','#bebada','#fb8072','#80b1d3','#fdb462','#fccde5')
plotly::plot_ly() %>%
  plotly::add_pie(data = pie_rel_to_island_100,
                  labels = ~ relation_to_island,
                  values = ~ n,
                  type = "pie",
                  #title = list(text = "bums",
                  #             position = "top right",
                  #             font = list(size = 40)),
                  textposition = "outside",
                  textinfo = "label+percent",
                  outsidetextfont = list(family = "Helvetica",
                                         size = 14),
                  marker = list(colors = colours,
                                line = list(color = '#FFFFFF', width = 3)),
                  showlegend = FALSE,
                  sort = TRUE,
                  direction = "clockwise",
                  domain = list(x = c(0, 0.35), y = c(0, 0.7))
                  ) %>%
  plotly::add_pie(data = pie_rel_to_island_all,
                  labels = ~ relation_to_island,
                  values = ~ n,
                  type = "pie",
                  #title = list(text = "bums",
                  #             position = "top right",
                  #             font = list(size = 40)),
                  textposition = "outside",
                  textinfo = "label+percent",
                  outsidetextfont = list(family = "Helvetica",
                                         size = 14),
                  marker = list(colors = colours,
                                line = list(color = '#FFFFFF', width = 3)),
                  showlegend = FALSE,
                  sort = TRUE,
                  direction = "clockwise",
                  domain = list(x = c(0.60, 0.95), y = c(0, 0.7))
                  )

```

### Refgene group

```{r, fig.cap="RefGene group of the top 100 most significant array hits (*left*) and all hits (right)", echo = FALSE}
colours <- list('#b3de69','#8dd3c7','#ffffb3','#bebada','#fb8072','#80b1d3','#fdb462','#fccde5')
plotly::plot_ly() %>%
  plotly::add_pie(data = pie_ref_group_100,
                  labels = ~ ucsc_refgene_group,
                  values = ~ n,
                  type = "pie",
                  #title = list(text = "bums",
                  #             position = "top right",
                  #             font = list(size = 40)),
                  textposition = "outside",
                  textinfo = "label+percent",
                  outsidetextfont = list(family = "Helvetica",
                                         size = 14),
                  marker = list(colors = colours,
                                line = list(color = '#FFFFFF', width = 3)),
                  showlegend = FALSE,
                  sort = TRUE,
                  direction = "clockwise",
                  domain = list(x = c(0, 0.35), y = c(0, 0.7))
                  ) %>%
  plotly::add_pie(data = pie_ref_group_all,
                  labels = ~ ucsc_refgene_group,
                  values = ~ n,
                  type = "pie",
                  #title = list(text = "bums",
                  #             position = "top right",
                  #             font = list(size = 40)),
                  textposition = "outside",
                  textinfo = "label+percent",
                  outsidetextfont = list(family = "Helvetica",
                                         size = 14),
                  marker = list(colors = colours,
                                line = list(color = '#FFFFFF', width = 3)),
                  showlegend = FALSE,
                  sort = TRUE,
                  direction = "clockwise",
                  domain = list(x = c(0.60, 0.95), y = c(0, 0.7))
                  )

```

# Session info

```{r}
sessionInfo()
```

<!-- ## Top 5 CpG regression diagnostics {.tabset .tabset-fade .tabset-pills} -->

<!-- ### CpG 1 -->

<!-- ```{r, echo=FALSE, message=FALSE, warning=FALSE} -->
<!-- plots[[3]][[1]] -->
<!-- ``` -->

<!-- ### CpG 2 -->

<!-- ```{r, echo=FALSE, message=FALSE, warning=FALSE} -->
<!-- plots[[3]][[2]] -->
<!-- ``` -->

<!-- ### CpG 3 -->

<!-- ```{r, echo=FALSE, message=FALSE, warning=FALSE} -->
<!-- plots[[3]][[3]] -->
<!-- ``` -->

<!-- ### CpG 4 -->

<!-- ```{r, echo=FALSE, message=FALSE, warning=FALSE} -->
<!-- plots[[3]][[4]] -->
<!-- ``` -->

<!-- ### CpG 5 -->

<!-- ```{r, echo=FALSE, message=FALSE, warning=FALSE} -->
<!-- plots[[3]][[5]] -->
<!-- ``` -->






<!-- # missMethyl {.tabset .tabset-pills} -->

<!-- ```{r missMethyl, include = FALSE} -->

<!-- Universe<-var$row.names -->
<!-- Sig_0.25<-var$row.names[var$adj.P.Val<0.25] -->
<!-- top100_named_genes <- var[!(is.na(var$ucsc_refgene_name) | var$ucsc_refgene_name == ""),] -->
<!-- top_100_cpgs <- top100_named_genes %>% arrange(adj.P.Val) %>%  dplyr::slice(1:100) %>% pull(row.names) -->

<!-- #FDR<0.25 -->
<!-- results_0.25_go <- epicR::mm_sig_cpgs(Sig_0.25) -->

<!-- ## 1:100 -->
<!-- #top_100_cpgs<-as.character(top_100_cpgs) Tests gene ontology enrichment for -->
<!-- #significant CpGs from Illumina's Infinium HumanMethylation450 or -->
<!-- #MethylationEPIC array, taking into account the differing number of probes per -->
<!-- #gene present on the array. -->
<!-- Universe<-as.character(Universe) -->
<!-- gst_top100_go<-gometh(sig.cpg=top_100_cpgs,  -->
<!--                       all.cpg = Universe,  -->
<!--                       collection = "GO",  -->
<!--                       array.type = "EPIC" -->
<!--                       ) -->
<!-- colnames(gst_top100_go)[2]<-"Term" -->

<!-- ### Results 1:100 - topGO pulls the most significant from the pack -->
<!-- results_top100_go <- topGO(gst_top100_go, number = 50) -->
<!-- results_top100_go <-  -->
<!--   results_top100_go %>%  -->
<!--   arrange(ONTOLOGY) -->
<!-- ``` -->

<!-- ## Significant hits (FDR<0.25) -->

<!-- ```{r miss methyl tables, echo = FALSE} -->
<!-- epicR::mm_plot_0.25(results_0.25_go) -->
<!-- ``` -->

<!-- ## Top 100 cpgs -->

<!-- ```{r, echo = FALSE} -->
<!-- results_top100_go %>%  -->
<!--   DT::datatable(filter = 'top', -->
<!--                 rownames = FALSE, -->
<!--                 extensions = 'FixedColumns', -->
<!--                 options = list( -->
<!--                   pageLength = 10,  -->
<!--                   dom = 'tpf', -->
<!--                   scrollX = TRUE, -->
<!--                   fixedColumns = TRUE -->
<!--                   ) -->
<!--                 ) %>%  -->
<!--   DT::formatStyle(columns = c("ONTOLOGY", "N", "DE", "P.DE", "FDR"),  -->
<!--                   `text-align` = 'center') %>% -->
<!--   DT::formatSignif(columns = c("P.DE", "FDR"), digits = 3)  -->
<!-- ``` -->
